Distributed Decision Making Model for Dynamic Formation of Web services Communities

### 2015-01-08 papers/webservice-coalition-learning/webservice_coalition_learning.tex
chnanges
changes

\begin{document}
%
% paper title
% can use linebreaks \\ within to get better formatting as desired
%\title{A Distributed Mechanism for Intelligent \\ Web Service Community Formation}
\title{DDM: Distributed Decision Making Model for Dynamic Formation of Web services Communities}
%
%
%
% author names and IEEE memberships
% note positions of commas and nonbreaking spaces ( ~ ) LaTeX will not break
% a structure at a ~ so this keeps an author's name from being broken across
% two lines.
% use \thanks{} to gain access to the first footnote area
% a separate \thanks must be used for each paragraph as LaTeX2e's \thanks
% was not built to handle multiple paragraphs
%
%
%\IEEEcompsocitemizethanks is a special \thanks that produces the bulleted
% lists the Computer Society journals use for "first footnote" author
% affiliations. Use \IEEEcompsocthanksitem which works much like \item
% for each affiliation group. When not in compsoc mode,
% \IEEEcompsocitemizethanks becomes like \thanks and
% \IEEEcompsocthanksitem becomes a line break with idention. This
% facilitates dual compilation, although admittedly the differences in the
% desired content of \author between the different types of papers makes a
% one-size-fits-all approach a daunting prospect. For instance, compsoc
% journal papers have the author affiliations above the "Manuscript
% received ..."  text while in non-compsoc journals this is reversed. Sigh.

@@

Web services are loosely-coupled business applications willing to cooperate in distributed settings within different groups called communities. Communities aim to provide web services with better visibility, efficiency, market share and total payoff. A number of mechanisms and models have been recently proposed to aggregate web
services and make them cooperate within their communities. However, the solutions mostly focus on short term goals. Moreover, these solutions generally suffer from high computational complexity, which makes decision making impossible in an on-demand fashion. In this paper, we propose a strategic decision making mechanism that regulates web service agents' decision making process in terms of cooperating with one another. This model first generates a trained set of data based on information obtained from large number of web services regarding their single and cooperative utilities as well as environmental parameters such as demand, service quality, etc. Utilizing the training model, the decision making mechanism implements a decision tree of possible viable strategies and their long-term expected utility gain.
%Utilizing the training model, the decision making mechanism implements a decision tree with an estimation of long-term utility gain for the web services. Communities can used the trained model and instantly choose best-response strategies considering their long-term gain. 
Our findings show that using this data, communities of web services can efficiently find the appropriate web service to invite cooperating with them as well as a single web service can find best communities to target joining them. In further details, after being trained, web services get to compute expectations as utilities they would gain while cooperating with communities of different characteristics. In contrast, communities can consider the training into account and analyze the characteristics of different individual web services and make prudent decisions while inviting a web service to join or accepting a join inquiry initiated from a web service. In general, DDM equips web services with efficient methods of foreseeing how their choices of actions would impact their long-term and short-term goals, therefore opting for best decision available. The experimental results show that our algorithms provide web services and community owners, in real-world-like environments, with applicable and near-perfect decision making mechanisms.

@@

In recent work communities of web services have been proposed that facilitate discovery of web services, improve the Quality of Service (QoS), and help individual services find better market share and opportunities. However, there is a twofold challenge that has still not been addressed. The first challenge is community development, which involves finding the right web services for the community, satisfying all web services and community coordinators involved within the community; and the second challenge is from the individual web service's side , such that a web service seeks communities from which to obtain the best outcome. A number of proposals have attempted to address these research problems. These solutions suffer from two main issues: 
\begin{enumerate}
	\item The solutions consider a centralized manager for communities and most of the decision making process for communities happens at a centralized entity. The problem is in real world scenarios most service providers work in distributed settings and having a centralized entity making decisions is not very feasible. 
	\item The solutions propose complex algorithms \cite{DBLP:conf/IEEEscc/LimTMB12, 10.1109/TSC.2012.12, 10.1109/TSC.2014.2312940} to find the optimal strategy, for \emph{each time} services or communities adopt an strategy. Or the problem is reduced by eliminating important parameters.
\end{enumerate}
Some of these works propose complex on-demand algorithms \cite{DBLP:conf/IEEEscc/LimTMB12, 10.1109/TSC.2012.12, 10.1109/TSC.2014.2312940} to effectively find target points, but they have to simplify the algorithm with limited parameters in order to make it feasible and algorithm executable. Simplified problems sometimes negatively influence the outcome because simplifying the constraints may cause important aspects of the problem to be ignored. For example, \cite{10.1109/TSC.2012.12} proposes a simple egalitarian way of distributing gain, which completely ignores gain for collaborative work of sub-communities instead of calculating the complex shapely-value gain distribution method. This model considers gain for all subsets of agents working together, which basically ignores the most important criteria of finding the optimal group. Other categories of related work exist that do not focus on all the parties involved within the community \cite{DBLP:conf/IEEEscc/KhosravifarABT11}. 
%And some other work which do not focus on rationality on web services or communities involved [refrences]. 
These solutions generally suffer from high complexity, which makes decision making impossible in an on-demand fashion, or they simplify important aspects to make it practical in the real world, thereby hurting the decision making performance.

In this paper, we introduce DDM (Distributed Learning-based Intelligent Service Agents) community framework that operates based on a decision model that regulates web service agents' decision making process in terms of cooperating with one another. The decision model in extracted from a data model in the form of information obtained from a large number of web services regarding their single and cooperative utilities as well as environmental parameters such as demand, service quality, etc. 
%The training model deploys a logistic regression algorithm to build a hypothesis function that can thoroughly address the aforementioned research problems. 
The decision model utilizes web services with enough information to efficiently decide and predict the outcome of their moves amongst the available services accessible to them. This model works in distributed manner in which our services are self-sufficient in decision making process and do not rely on a centralized decision maker process. Our findings show that communities of web services can efficiently find the appropriate web service to invite cooperating with them as well as allowing a single web service to find the best communities to join them. Furthermore, the decision model provides web services with expectations of utilities they would gain while cooperating with communities of different characteristics. In contrast, communities can consider the decision model and analyze the characteristics of different individual web services and make prudent decisions when inviting a web service to join or accepting a join inquiry initiated from a web service. In general, DDM equips web services with efficient methods for foreseeing how their choices will impact their long-term and short-term goals; therefore, opting for best decision available.

@@

The rest of the paper is organized as follows. In Section 2 we briefly describe the preliminaries and problem statement, discusses the web service community architecture, and the considered parameters in our feature vectors. Section 3 presents our solution model, the way data is trained and web services utilize them in their decision making process. Section 4 describes our experiments and results. Section 5 discussed relevant related work. Finally Section 6, concludes the paper and identifies future work.

@@

\section{Preliminaries and Problem Statement}\label{s:preliminaries}
In this section, we first present the architecture of DDM. We explore the characteristics of intelligent service agents and the features we extract for training. To do this, we first discuss some preliminaries.

@@

@@

\subsubsection{Feature Normalization}\label{sss:fn}

@@

\section{The Model}\label{s:model}

@@

\subsubsection{web services}\label{sss:webservices}

@@

\subsubsection{filtered data}\label{sss:filtereddata}

@@

\subsubsection{feature engineering}\label{sss:feng}
Let $CFVS = \{C_1, C_2,..., C_N\}$ be the community feature vector set with $N$ communities. Based on the $CFVS$ set, we create an $|N \times N|$ gain matrix for $t$ different time slots. Each entry of $gain_{n,m}^{t}$corresponds to a utility gain of web services in $C_n$, when $C_n$ joins $C_m$, based on community and web service data available for time $t$. All of these values have to be evaluated by implementing the community with members of $ws_i \in C_n$ and $ws_j \in C_m$ and evaluating its utility (see Equation \ref{u_c_load}). Therefore $gain_{n,m}^{t} = U_{C_n \cup C_m}^{t} - U_{C_{n}}^{t}$. Evaluating the utility gain for all entries of the $gain$ matrix can be a costly process with a very big number of $N$, size of the feature vector $|N|$ should be chosen carefully.

@@

When communities $C_i$ and $C_j$ are in each other's top $k$ preference set, their combination is added to the list of possible communities that can join others at time $t+1$. Also, for each community $C_i$ in our initial $CFVS$ set, we maintain a tree with the community $C_i$ as its root. Its children are all of the communities that $C_i$ decided to join. As the scenario progresses over time, the merged communities may decide to join other communities. When community $C_i$ and $C_j$ decide to join each other and create community $C_k$, the new community $C_k$ will be added as a child to both $C_i$ and $C_j$ nodes. At the end of the process, each community is utilized with a tree of set of possible combination of communities it can join.

@@

\subsection{Decision Tree Creation Complexity}\label{ss:complexity}
Here we analyze the computational complexity of the DDM decision tree creation algorithm on each time iteration $t$. Computing top $k$ preferred communities for $C_i$ in $L^t(C_i, k)$ requires an $O(n.log(n))$ sort time. $C_i$ has a result set of size $|k|$, and for each of those $k$ communities we need to check against their $k$ top preferred communities. Considering we already have the list sorted, line 6 in algorithm runs $O(k \times n.log(n))$ times. Doing this search for all communities in $|CFVS|$ set $n$ times (line 5), the overall order of complexity of the algorithm in regards with $n$ and $k$ is: $O(k^2 \times n^2.log(n))$.

>>>>

\begin{document}
%
% paper title
% can use linebreaks \\ within to get better formatting as desired
%\title{A Distributed Mechanism for Intelligent \\ Web Service Community Formation}
\title{Distributed Decision Making Model for Dynamic Formation of Web services Communities}
%
%
%
% author names and IEEE memberships
% note positions of commas and nonbreaking spaces ( ~ ) LaTeX will not break
% a structure at a ~ so this keeps an author's name from being broken across
% two lines.
% use \thanks{} to gain access to the first footnote area
% a separate \thanks must be used for each paragraph as LaTeX2e's \thanks
% was not built to handle multiple paragraphs
%
%
%\IEEEcompsocitemizethanks is a special \thanks that produces the bulleted
% lists the Computer Society journals use for "first footnote" author
% affiliations. Use \IEEEcompsocthanksitem which works much like \item
% for each affiliation group. When not in compsoc mode,
% \IEEEcompsocitemizethanks becomes like \thanks and
% \IEEEcompsocthanksitem becomes a line break with idention. This
% facilitates dual compilation, although admittedly the differences in the
% desired content of \author between the different types of papers makes a
% one-size-fits-all approach a daunting prospect. For instance, compsoc
% journal papers have the author affiliations above the "Manuscript
% received ..."  text while in non-compsoc journals this is reversed. Sigh.

@@

Web services are loosely-coupled business applications willing to cooperate in distributed settings within different groups called communities. Communities aim to provide web services with better visibility, efficiency, market share and total payoff. A number of mechanisms and models have been recently proposed to aggregate web
services and make them cooperate within their communities. However, the solutions mostly focus on short term goals. Moreover, these solutions generally suffer from high computational complexity, which makes decision making impossible in an real-time fashion. In this paper, we propose a strategic distributed decision making mechanism called DDM that regulates web service agents' decision making process in terms of cooperating with one another. This model first generates a trained set of data based on information obtained from large number of web services regarding their single and cooperative utilities as well as environmental parameters such as demand, service quality, etc. Utilizing the training model, the decision making mechanism implements a decision tree of possible viable strategies and their long-term expected utility gain.
%Utilizing the training model, the decision making mechanism implements a decision tree with an estimation of long-term utility gain for the web services. Communities can used the trained model and instantly choose best-response strategies considering their long-term gain. 
Our findings show that using this data, communities of web services can efficiently find the appropriate web service to invite cooperating with them as well as a single web service can find best communities to target joining them. In further details, after being trained, web services get to compute expectations as utilities they would gain while cooperating with communities of different characteristics. In contrast, communities can consider the training into account and analyze the characteristics of different individual web services and make prudent decisions while inviting a web service to join or accepting a join inquiry initiated from a web service. In general, DDM equips web services with efficient methods of foreseeing how their choices of actions would impact their long-term and short-term goals, therefore opting for best decision available. The experimental results show that our algorithms provide web services and community owners, in real-world-like environments, with applicable and near-perfect decision making mechanisms.

@@

\textbf{Challenges and Problem Statement.} In recent work communities of web services have been proposed that facilitate discovery of web services, improve the Quality of Service (QoS), and help individual services find better market share and opportunities. However, there is a twofold challenge that has still not been addressed. The first challenge is community development, which involves finding the right web services for the community, satisfying all web services and community coordinators involved within the community; and the second challenge is from the individual web service's side , such that a web service seeks communities from which to obtain the best outcome. A number of proposals have attempted to address these research problems. These solutions suffer from two main issues: 
\begin{enumerate}
	\item The solutions consider a centralized manager for communities and most of the decision making process for communities happens at a centralized entity. The problem is in real world scenarios most service providers work in distributed settings and having a centralized entity making decisions is not very feasible. 
	\item The solutions propose complex algorithms \cite{DBLP:conf/IEEEscc/LimTMB12, 10.1109/TSC.2012.12, 10.1109/TSC.2014.2312940} to find the optimal strategy, for \emph{each time} services or communities adopt an strategy. Or the problem is reduced by eliminating important parameters.
\end{enumerate}
Some of these proposals have complex algorithms \cite{DBLP:conf/IEEEscc/LimTMB12, 10.1109/TSC.2012.12, 10.1109/TSC.2014.2312940} to effectively find target points, but they have to reduce the complexity of the algorithms using simplified approximation algorithms in order to make it feasible and algorithm executable in real-time. These approximation methods sometimes negatively influence the outcome because simplifying the constraints may cause important aspects of the problem to be ignored. For example, \cite{10.1109/TSC.2012.12} proposes a simple egalitarian way of distributing gain, which completely ignores gain for collaborative work of sub-communities instead of calculating the complex shapely-value gain distribution method. This model considers gain for all subsets of agents working together, which basically ignores the most important criteria of finding the optimal group. Other categories of related work exist that do not focus on all the parties involved within the community \cite{DBLP:conf/IEEEscc/KhosravifarABT11}. 
%And some other work which do not focus on rationality on web services or communities involved [refrences]. 
These solutions generally suffer from high complexity, which makes decision making impossible in an real-time fashion, or they simplify important aspects to make it practical in the real world, thereby hurting the decision making performance. In \cite{journal-community-formation} we proposed a cooperative game theory based model for aggregating the web services in communities. A centralized decision maker in communities based on a complete knowledge of available web service quality metrics and performance, forms optimal and stable communities to maximize individual and group efficiency and income.

\textbf{Contributions.} In this paper, we introduce DDM, a community framework that operates based on a decision model that regulates web service agents' decision making process in terms of cooperating with one another. The decision model in extracted from a data model in the form of information obtained from a large number of web services regarding their single and cooperative utilities as well as environmental parameters such as demand, service quality, etc. 
%The training model deploys a logistic regression algorithm to build a hypothesis function that can thoroughly address the aforementioned research problems. 
The decision model utilizes web services with enough information to efficiently decide and predict the outcome of their moves amongst the available services accessible to them. This model works in distributed manner in which our services are self-sufficient in decision making process and do not rely on a centralized decision maker process. Our findings show that communities of web services can efficiently find the appropriate web service to invite cooperating with them as well as allowing a single web service to find the best communities to join them. Furthermore, the decision model provides web services with expectations of utilities they would gain while cooperating with communities of different characteristics. In contrast, communities can consider the decision model and analyze the characteristics of different individual web services and make prudent decisions when inviting a web service to join or accepting a join inquiry initiated from a web service. In general, DDM equips web services with efficient methods for foreseeing how their choices will impact their long-term and short-term goals; therefore, opting for best decision available.

@@

\textbf{Organization.} The rest of the paper is organized as follows. In Section 2 we briefly describe the preliminaries and problem statement, discusses the web service community architecture, and the considered parameters in our feature vectors. Section 3 presents our solution model, the way data is trained and web services utilize them in their decision making process. Section 4 describes our experiments and results. Section 5 discussed relevant related work. Finally Section 6, concludes the paper and identifies future work.

@@

\section{Preliminaries and Challenges}\label{s:preliminaries}
%In this section, we first present the architecture of DDM. We explore the characteristics of intelligent service agents and the features we extract for training. To do this, we first discuss some preliminaries.
In this section, we discuss the preliminary concepts of communities of web service and we introduce the challenges of community formation.

@@

\section{The Model Components}\label{s:themodelcomponents}

In this section, we discuss the parameters that we use in the rest of the paper. Then we present the task distribution and revenue model of our distributed web services communities.

@@

@@

\section{The Decision Making Mechanism}\label{s:model}

@@

\subsubsection{Web Service data}\label{sss:webservices}

@@

\subsubsection{Filtered Data}\label{sss:filtereddata}

@@

\subsubsection{Feature Engineering}\label{sss:feng}
Let $CFVS = \{C_1, C_2,..., C_N\}$ be the community feature vector set with $N$ communities. Based on the $CFVS$ set, we create an $|N \times N|$ gain matrix for $t$ different time slots. Each entry of $gain_{n,m}^{t}$corresponds to a utility gain of web services in $C_n$, when $C_n$ joins $C_m$, based on community and web service data available for time $t$. All of these values have to be evaluated by implementing the community with members of $ws_i \in C_n$ and $ws_j \in C_m$ and evaluating its utility (see Equation \ref{u_c_load}). Therefore $gain_{n,m}^{t} = U_{C_n \cup C_m}^{t} - U_{C_{n}}^{t}$. Evaluating the utility gain for all entries of the $gain$ matrix can be a costly process with a very big number of $N$, size of the feature vector $|N|$ should be chosen carefully.

@@

When communities $C_i$ and $C_j$ are in each other's top $k$ preference set, their combination is added to the list of possible communities that can join others at time $t+1$. Also, for each community $C_i$ in our initial $CFVS$ set, we maintain a tree with the community $C_i$ as its root. Its children are all of the communities that $C_i$ decided to join. As the scenario progresses over time, the merged communities may decide to join other communities. When community $C_i$ and $C_j$ decide to join each other and create community $C_k$, the new community $C_k$ will be added as a child to both $C_i$ and $C_j$ nodes. At the end of the process, each community is utilized with a tree of set of possible combination of communities it can join. Algorithm \ref{algo:dectree} illustrates the DDM tree creation procedure as pseudo-code.

@@

\textbf{Complexity.} Here we analyze the computational complexity of the DDM decision tree creation algorithm on each time iteration $t$. Computing top $k$ preferred communities for $C_i$ in $L^t(C_i, k)$ requires an $O(n.log(n))$ sort time. $C_i$ has a result set of size $|k|$, and for each of those $k$ communities we need to check against their $k$ top preferred communities. Considering we already have the list sorted, line 6 in algorithm runs $O(k \times n.log(n))$ times. Doing this search for all communities in $|CFVS|$ set $n$ times (line 5), the overall order of complexity of the algorithm in regards with $n$ and $k$ is: $O(k^2 \times n^2.log(n))$.
